---
title: "ESM244 Session 3 (Make-up Lecture)"
author: "Anna Abelman"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# 1. Read in data
library(tidyverse)
library(janitor)
library(sf)
library(tmap)
```

#### 2. Read in CA eco-regions data

```{r}
ca_eco <- read_sf(dsn = ".", layer = "ca_eco") %>%
  select(US_L3NAME) %>% 
  rename(region = US_L3NAME) %>% 
  st_simplify(dTolerance = 100) %>% #simplifying the polygon; be careful not to oversimplify 
  st_transform(4326) #setting the EPSG
 
#rmapshaper::ms_simplify(), geom::gSimplify are other options

# to check by using st_crs(ca_eco) to set EPSG

plot(ca_eco)

```

#### 3. Read in CA COunties polygons

```{r}
ca_counties <- read_sf(dsn = ".", layer = "california_county_shape_file")
#check using st_crs; originally says NA, so change to 4326
st_crs(ca_counties) = 4326
```

#### 4. Read in CA dams location and data
```{r}
ca_dams <- read_sf(dsn = ".", layer = "California_Jurisdictional_Dams")
#check st_crs; already set to 4326
```

#### 5. Make a map of dams in CA, overlayed on eco-regions, with outlines of CA counties

```{r}

ggplot(data = ca_counties)+
  geom_sf(color = "black",
          size = 0.1)+
  geom_sf(data = ca_eco,
          aes(fill = region),
          alpha = 0.5,
          color = "NA",
          show.legend = FALSE)+
  geom_sf(data = ca_dams,
          size = 0.5,
          alpha = 0.5)+
  theme_minimal()

```

#### 6. Just explore SB county eco-regions

```{r}

sb <- ca_counties %>% 
  filter(NAME == "Santa Barbara")

#clip eco-region data to only include information within SB county - use st_interection

eco_clip <- st_intersection(ca_eco, sb)

#now let's plot it:

ggplot()+
  geom_sf(data = ca_counties,
          fill = "gray90",
          color = "gray80",
          size = 0.2)+
  geom_sf(data = eco_clip,
          aes(fill = region),
          color = "white",
          show.legend = FALSE)+
  coord_sf(xlim = c(-121,-119), 
           ylim = c(33.5,35.5))+ # crops to only show a small portion of CA
  scale_fill_manual(values = c("purple", "gold2", "blue"))

```

#### 7. Create an interactive map with tmap

```{r}
#use link in server to find different basemaps
sb_clip_tmap <- tm_basemap("Esri.WorldImagery")+
  tm_shape(eco_clip)+
  tm_fill("region", palette = c("orange", "green", "purple"),
          alpha = 0.5)

tmap_mode("view")
sb_clip_tmap
```

#### 8. How to create an `sf` object from latitude/longitude recordings (e.g. in an Excel file)

Make a mock dataset using tibble():
```{r}

my_example <- tribble(
  ~id, ~lon, ~lat,
  "tiger", -119.4, 34.35,
  "lion", -119.41, 34.39,
  "bear", -119.43, 34.38
)
#need R to recognize as spatial information to be able to plot using tmap

#convert this to `sf` object:
animals_sf <- st_as_sf(my_example, coords = c("lon", "lat"), crs = 4326)

```

Make a map using tmap
```{r}

animal_map <- tm_basemap("Stamen.Watercolor")+
  tm_shape(animals_sf)+
  tm_dots(labels = "id", col = "skyblue",
          size = 0.5)

animal_map
```

#### 9. Chloropleth of dam counts

```{r}
intersection <- st_intersection(x = ca_dams, y = ca_counties)

dams_per_county <- intersection %>% 
  group_by(NAME) %>% 
  tally()

ca_tot <- ca_counties %>%
  st_join(dams_per_county) %>% 
  select(NAME.x, n) %>% 
  rename(name = NAME.x)

ca_tot_zero <- ca_tot %>% 
  replace_na(list(n = 0))

#one more example with ggplot

ggplot()+
  geom_sf(data = ca_tot_zero, 
          aes(fill = n),
          size = 0.2,
          color = "white")+
  scale_fill_continuous(low = "yellow", high = "red")

```

















